---
title: "dplyr Package"
author: "23BMAT019"
date: "2023-10-20"
output: 
  html_document: 
    toc: yes
    highlight: kate
    theme: spacelab
---
# Adding `dplyr`
```{r}
#library("dplyr")
library(hflights)
```

# Chaining and Pipelining
**Syntax**: x `%>%` f() `# means f(x)`

This process is called chaining. 

```{r}
# %>% pipeline symbol: 
a <- 1:5
sum(a)
#a %>% sum()

# sqrt(sum((x1-x2)^2)), where x1 and x2 are vectors.
x1 <- 1:5
x2 <- x1 + 1

sqrt(sum((x1-x2)^2))
#(x1-x2)^2 %>% sum() %>% sqrt()
```

# `dplyr` functionality

Five Basic verbs: `filter`, `select`, `arrange`, `mutate`, and `summarise`

## Extract Cases

### `filter`: keep rows satisfying a specific condition
**Syntax**: `filter`(`data`, `condition`, ...)

```{r}
#1 Find the rows where Sepal.Length > 5
filter(iris, Sepal.Length > 5) # `$` is not needed

#2 Find the rows where Sepal.Length > 4 and Species is setosa
filter(iris, Sepal.Length > 4 & Species=="setosa")

#3 Find the rows where Sepal.Length > 2 or Species is setosa
filter(iris, Sepal.Length > 2 | Species=="setosa")

#4 Find the rows where Species is either setosa or versicolor
filter(iris, Species=="setosa" |  Species=="versicolor")
filter(iris, Species %in% c("setosa", "versicolor")) # Alternative
```

```{r}
View(hflights)

#1 Find all the flights on Jan 1

## with dplyr
### without chaining
filter(hfilghts, Month==1 & DayofMonth==1)
### with chaining [HW]

## without dplyr
### without attach/detach
hflights[hflights$Month==1 & hflights$DayofMonth==1,]
### without attach/detach
attach(hflights)
hflights[Month==1 & DayofMonth==1,]
detach(hflights)

#2 All flights of UniqueCarrier AA or UA

## with dplyr
filter(hflights, UniqueCarrier %in% c("AA", "UA"))
## without dplyr
hflights[hflights$UniqueCarrier %in% c("AA", "UA"), ]
```

## Extract Variables

### `select`: pick column by name
**Syntax**: `select`(`data`, `col1`, `col2`, ...)

```{r}
#1 Extract the data for Sepal.Length and Petal.Length
select(iris, Sepal.Length, Petal.Length)

#2 Extract the column which start with Petal
select(iris, starts_with("Petal"))

#3 Extract the column which end with Length
select(iris, ends_with("Length"))

#4 Extract the column which start with Sepal and contains Species.
select(iris, starts_with("Sepal"), Species) # Logically incorrect
select(iris, starts_with("Sepal"), contains("Species"))

#5 Extract the column which start with Sepal and contains Length.
select(iris, starts_with("Sepal"), contains("Length")) # 3 columns

#6 Select DepTime, ArrTime and FlightNum columns
## with dplyr
select(hflights, DepTime, ArrTime, FlightNum)
## without dplyr
hflights[,c("DepTime", "ArrTime", "FlightNum")]  # using [,col]
hflights[c("DepTime", "ArrTime", "FlightNum")]   # using [col]

#7 Select multiple continuous columns
# i. from DayofMonth to FlightNum
select(hflights, DayofMonth:FlightNum)

# ii. from Year to ArrTime and alongwith columns that contain "Taxi" and "Delay"
select(hflight, Year:ArrTime, contains("Taxi"), contains("Delay"))
```

```{r}
# CQ1. Select UniqueCarrier and DepDelay Column and filter for Departure delays over 60 minutes
filter(select(hflights, UniqueCarrier, DepDelay), DepDelay > 60)
## find dimensions
dim(filter(select(hflights, UniqueCarrier, DepDelay), DepDelay > 60))

## use chaining
hflights %>% select(UniqueCarrier, DepDelay) %>% filter(DepDelay > 60)
hflights %>% select(UniqueCarrier, DepDelay) %>% filter(DepDelay > 60) %>% dim()
```

## Arranging

### `arrange`: To reorder the rows
**Syntax**: `arrange`(`data`, `parameter`)

```{r}
#1 Select  UniqueCarrier and DepDelay columns and sort by DepDelay
select(hflights, UniqueCarrier, DepDelay)
arrange(select(hflights, UniqueCarrier, DepDelay), DepDelay)
## Chaining Counterpart
hflights %>% select(UniqueCarrier, DepDelay) %>% arrange(DepDelay)
## Descending
hflights %>% select(UniqueCarrier, DepDelay) %>% arrange(desc(DepDelay))
```
