---
title: "Practice"
author: "You"
date: "2023-11-07"
output: html_document
---

```{r}
# Add packages to library
library(readxl)
library(openxlsx)
```

# Functions with Syntax

1. `write`.`xlsx`(`dataframe`, "`filename.xlsx`")
  
  - Writes the `dataframe` to a file named `filename.xlsx`.
  - A Workbook object is returned
  - Please note that this function is only for .xlsx files

2. `read`.`xlsx`("`filename.xlsx`")

  - Reads `filename.xlsx` 
  - A data frame is returned

**Note:**
`read`  - **Extract** data
`write` - **Modify**  data

# Converting a Data Frame to Excel Workbook

```{r}
getwd()

head(mtcars)
class(mtcars)

# To store a Workbook object in cars.
cars <- write.xlsx(mtcars, "cars.xlsx")
## The workbook object converts `mtcars` 
## into an Excel Workbook named `cars.xlsx`

cars # Workbook object

my_wbook <- write.xlsx(iris, "23bmat019.xlsx")
my_wbook
```

# (Reading Files into R) Converting an Excel Workbook to a Data Frame 

```{r}
my_df <- read.xlsx("cars.xlsx")
my_df
```


